---
title: "Ohio"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, include=FALSE}
library(readstata13)
library(dplyr)
library(ggplot2)
library(tidyverse)
```
# Analysis of Ohio Education and Tractor Data

```{r data-in, echo = FALSE, warning = FALSE}
# This 1940 Edu dataset is already filtered to White Males 18-22
edu1940 <- read.csv("Edu_Raw_Data/OhioEduData/ohio1940edu.csv")
edu1940[,1] <- NULL
edu1930 <- read.csv("Edu_Raw_Data/OhioEduData/ohio1930edu.csv")
edu1930 <- subset(edu1930, sex == "Male")
edu1930 <- subset(edu1930, between(age, 18, 22))
edu1930[,1] <- NULL

ohio1930Tractor <- read_csv("Tractor_Raw_Data/OhioTractorData/ohioTractor1930.csv")
ohio1930Tractor$county <- ohio1930Tractor$county * 10
#Remove Pickaway
ohio1930Tractor <- subset(ohio1930Tractor, county != 1290)
ohio1930Tractor[,c(1,2)] <- NULL


ohio1940Tractor <-read_csv("Tractor_Raw_Data/OhioTractorData/ohioTractor1940.csv")
ohio1940Tractor$county <- ohio1940Tractor$county * 10
#Remove Pickaway
ohio1940Tractor <- subset(ohio1940Tractor, county != 1290)
ohio1940Tractor[,c(1,2)] <- NULL


edu1940Avg <- edu1940 %>%
  group_by(county) %>%
  summarise(avg = mean(higrade))

df_40Edu30Tractor<- merge(edu1940Avg, ohio1930Tractor, by = "county")
rownames(df_40Edu30Tractor) <- 1:nrow(df_40Edu30Tractor)


#vam-ohio
vam.ohio <- read.csv("vamOhio.csv")
vam.ohio1947 <- subset(vam.ohio, year == 1947)
vam.ohio1954 <- subset(vam.ohio, year == 1954)
```

## Exploratory Plots of Average Education by County Vs. % of Farms with Tractors

### Ohio 1940 Average Education Vs. Ohio 1930 % of Farms with Tractor
```{r 1940EduVs1930Tractor, echo=FALSE}
low_counties <- df_40Edu30Tractor[c(65, 70,49, 24),]
close_counties <- df_40Edu30Tractor[c(63,35),]

ggplot(df_40Edu30Tractor, aes(x = percent_farm_tractor, y = avg)) + geom_point() + geom_path(data = low_counties, color = "red") + geom_point(data = close_counties, shape = 8, color = "red") + xlab("1930 % Farms with Tractor") + ylab("1940 Average Education") + ggtitle("Ohio Counties")
```



### Linear Model 1940 Avg. Edu vs 1930 % Farm w/ Tractor
```{r}
ohEduModLinear <- lm(avg ~ percent_farm_tractor, data = df_40Edu30Tractor)
plot(df_40Edu30Tractor$percent_farm_tractor, df_40Edu30Tractor$avg, main = "Linear")
abline(ohEduModLinear, col = "red")
summary(ohEduModLinear)
```


### Log-Linear Model 1940 Avg. Edu vs 1930 % Farm w/ Tractor
```{r}
df_40Edu30Tractor$logAvg <- log(df_40Edu30Tractor$avg)
ohEduModLogLinear <- lm(logAvg ~ percent_farm_tractor, data = df_40Edu30Tractor)
plot(df_40Edu30Tractor$percent_farm_tractor, df_40Edu30Tractor$logAvg, main = "Log-Linear")
abline(ohEduModLogLinear, col = "red")
summary(ohEduModLogLinear)

```

### Linear-Log Model 1940 Avg. Edu vs 1930 % Farm w/ Tractor
```{r}
df_40Edu30Tractor$logPercent <- log(df_40Edu30Tractor$percent_farm_tractor)
ohEduModLinearLog <- lm(avg ~ logPercent, data = df_40Edu30Tractor)
plot(df_40Edu30Tractor$logPercent, df_40Edu30Tractor$avg, main = "Linear-Log")
abline(ohEduModLinearLog, col = "red")
summary(ohEduModLinearLog)
```



```{r, echo=FALSE}
#Create VAM PlotSrc data (1947)
vam1947Tractor30 <- merge(df_40Edu30Tractor, vam.ohio1947, by = "fips")
vam1947Tractor30 <- vam1947Tractor30[!is.na(vam1947Tractor30$VAM),]
```




### Linear Model VAM 1947 vs Percent 1930
```{r}
vam47Tractor30Mod <- lm(VAM ~ percent_farm_tractor, data = vam1947Tractor30)
plot(vam1947Tractor30$percent_farm_tractor, vam1947Tractor30$VAM, main = "Linear")
abline(vam47Tractor30Mod, col = "red")
summary(vam47Tractor30Mod)
```


# Residuals from VAM 47 vs. Percent 30
```{r}
vam47Residuals <-data.frame(vam47Tractor30Mod$residuals, vam1947Tractor30$county, vam1947Tractor30)
vam47Residuals
```







```{r, echo=FALSE}
#Create VAM PlotSrc data (1954)
vam1954Tractor30 <- merge(df_40Edu30Tractor, vam.ohio1954, by = "fips")
vam1954Tractor30 <- vam1954Tractor30[!is.na(vam1954Tractor30$VAM),]
```




### Linear Model VAM 1947 vs Percent 1930
```{r}
vam54Tractor30Mod <- lm(VAM ~ percent_farm_tractor, data = vam1954Tractor30)
plot(vam1954Tractor30$percent_farm_tractor, vam1954Tractor30$VAM, main = "Linear")
abline(vam54Tractor30Mod, col = "red")
summary(vam54Tractor30Mod)
```


# Residuals from VAM 47 vs. Percent 30
```{r}
vam54Residuals <-data.frame(vam54Tractor30Mod$residuals, vam1954Tractor30$county, vam1954Tractor30)
vam54Residuals
```




# High Education and VAM(1947, 1954)

##1947
```{r}
plot(vam1947Tractor30$VAM, vam1947Tractor30$avg, main = "Education vs. VAM (1947)")
```






##1954
```{r}
plot(vam1954Tractor30$VAM, vam1954Tractor30$avg, main = "Education vs. VAM (1954)")
```













Notice the lowest points going across the X-Axis(On Plot with line connecting outlier counties). These counties identified left-to-right are:

* Pike County
* Ross County
* Madison County
* Fayette County

```{r low_counties_write_df, include = FALSE}
pike <- subset(edu1940, county == 1310)
# write.csv(pike, "pike_county.csv")
 ross <- subset(edu1940, county == 1410)
# write.csv(ross, "ross_county.csv")
 madison <- subset(edu1940, county == 970)
# write.csv(madison, "madison_county.csv")
 fayette <- subset(edu1940, county == 470)
# write.csv(fayette, "fayette_county.csv")
```

The 2 Counties which are overlaying each other and can be seen to the right of the outlier line are

* Paulding County
* Henry County

```{r overlap_counties_df, include = FALSE}
# paulding <- subset(edu1940, county == 1250)
# write.csv(paulding, "paulding_county.csv")
# henry <- subset(edu1940, county == 690)
# write.csv(henry, "henry_county.csv")
```


The objective was to look at the difference in education from 1930 and 1940 and plot that against the % of Farms with Tractors in 1930. 
The problem I encountered was that the 1930's Education dataset has not features which indicate the highest grade completed for an individual.

Taking a closer look at the outlier counties listed above(Pike, Ross, Madison, Fayette)
but removing everyone who is currently in school, as they could continue their education and in that case the current represented value would not be accurate.

## Highest Grade Completed by Age for the 4 outlier counties
```{r}
pike <- subset(pike, school == "No, not in school")
ggplot(pike, aes(x = age, y = higrade, group = age)) + geom_boxplot() + xlab("Age Groups") + ylab("Highest Grade Completed") + ggtitle("Pike County")
```

```{r}
ross <- subset(ross, school == "No, not in school")
ggplot(ross, aes(x = age, y = higrade, group = age)) + geom_boxplot() + xlab("Age Groups") + ylab("Highest Grade Completed") + ggtitle("Ross County")
```

```{r}
madison <- subset(madison, school == "No, not in school")
ggplot(madison, aes(x = age, y = higrade, group = age)) + geom_boxplot() + xlab("Age Groups") + ylab("Highest Grade Completed") + ggtitle("Madison County")
```

```{r}
fayette <- subset(fayette, school == "No, not in school")
ggplot(fayette, aes(x = age, y = higrade, group = age)) + geom_boxplot() + xlab("Age Groups") + ylab("Highest Grade Completed") + ggtitle("Fayette County")
```

