---
title: "ScatterPlots"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readstata13)
library(dplyr)
library(ggplot2)
```

Tractor & VAM Plots:  
  [Ohio](#ohio-tractor-vam-plots)
  


Note about VAM Plots: Any counties with `N/A` value for VAM were replaced with 0. These are so few that I believe it has minimal effect on the trend lines. 

```{r}
#Function for producing EDU dataset
#
#
eduDataFunc <- function(state){
  edu_data <- subset(oth_state, stateicp == state)
  edu_data <- subset(edu_data, sex == "Male")
  edu_data <- subset(edu_data, between(age, 18,22))
  edu_data <- subset(edu_data, higrade < 20)
  edu_data <- subset(edu_data, higrade > 0)
  edu_data$higrade <- edu_data$higrade - 3
  
  reduced_edu_data <- edu_data %>%
    group_by(county) %>%
    summarise(avg = mean(higrade))
  
  return(reduced_edu_data)
}
#
#
#

eduMedianFunc <- function(state){
  edu_data <- subset(oth_state, stateicp == state)
  edu_data <- subset(edu_data, sex == "Male")
  edu_data <- subset(edu_data, between(age, 18,22))
  edu_data <- subset(edu_data, higrade < 20)
  edu_data <- subset(edu_data, higrade > 0)
  edu_data$higrade <- edu_data$higrade - 3
  
  reduced_edu_data <- edu_data %>%
    group_by(county) %>%
    summarise(med = median(higrade))
  
  return(reduced_edu_data)
}


#Function for producing Tractor set given State Abb
#
#
tractDataFunc <- function(stateAbbre){
  stateDf <- subset(df30, stateAbb == stateAbbre)
  stateDf$county <- stateDf$county * 10
  stateDf$X <- NULL
  return(stateDf)
}
#
#
#


```


```{r, echo = FALSE}
df <- read.csv("Tractor_Raw_Data/df29-69.csv")
filtered <- read.csv("Tractor_Raw_Data/filtered_df.csv")
stateLvl <- read.csv("Tractor_Raw_Data/rolledUp_all.csv")
vam <- read.csv("vamFipsData.csv")
vam_1958 <- subset(vam, year == 1958)
vam_1963 <- subset(vam, year == 1963)
df25 <- subset(df, year == 1925)
df30<- subset(df, year == 1930)
df40<- subset(df, year == 1940)
df50<- subset(df, year == 1950)
df54<- subset(df, year == 1954)
df64<- subset(df, year == 1964)
df69<- subset(df, year == 1969)
```

---

#National Level

## % Farm Tractor 1940 vs 1930
```{r}
df30_40 <- merge(df30, df40, by = c("name", "stateAbb"), all = FALSE)
ggplot(data = df30_40, aes(x=percent_farm_tractor.x, y=percent_farm_tractor.y)) + geom_point() + xlab(label = "% of Farms w/ Tractors 1930's") + ylab(label = "% of Farms w/ Tractors 1940's") + ggtitle(label = "National Level")

```

## % Farm Tractor 1950 vs 1940
```{r}
df40_50 <- merge(df40, df50, by = c("name", "stateAbb"), all = FALSE)
ggplot(data = df40_50, aes(x=percent_farm_tractor.x, y=percent_farm_tractor.y)) + geom_point() + xlab(label = "% of Farms w/ Tractors 1940's") + ylab(label = "% of Farms w/ Tractors 1950's") + ggtitle(label = "National Level")
```


## % Farm Tractor 1954 vs 1950
```{r}
df50_54 <- merge(df50, df54, by = c("name", "stateAbb"), all = FALSE)
ggplot(data = df50_54, aes(x=percent_farm_tractor.x, y=percent_farm_tractor.y)) + geom_point() + xlab(label = "% of Farms w/ Tractors 1950") + ylab(label = "% of Farms w/ Tractors 1954") + ggtitle(label = "National Level")

df50_54$logResponse <- log(df50_54$percent_farm_tractor.y)
df50_54$logX <- log(df50_54$percent_farm_tractor.x)
ggplot(data = df50_54, aes(x=percent_farm_tractor.x, y=logResponse)) + geom_point() + xlab(label = "% of Farms w/ Tractors 1950") + ylab(label = "Log(% of Farms w/ Tractors 1954)") + ggtitle(label = "National Level") 
ggplot(data = df50_54, aes(x=logX, y=logResponse)) + geom_point() + xlab(label = "Log(% of Farms w/ Tractors 1950)") + ylab(label = "Log(% of Farms w/ Tractors 1954)") + ggtitle(label = "National Level")
```



## % Farm Tractor 1964 vs 1954
```{r}
df54_64 <- merge(df54, df64, by = c("name", "stateAbb"), all = FALSE)
ggplot(data = df54_64, aes(x=percent_farm_tractor.x, y=percent_farm_tractor.y)) + geom_point() + xlab(label = "% of Farms w/ Tractors 1954") + ylab(label = "% of Farms w/ Tractors 1964") + ggtitle(label = "National Level")
```


------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

#<a id="ohio-tractor-vam-plots"></a>OHIO

```{r}

edu <- read.dta13("ohio1940.dta")

ohio1930 <- subset(df30, stateAbb == "OH")
#Adjust FIPS code representation to match the EDU dataset
ohio1930$county <- ohio1930$county * 10
#Remove Pickaway
ohio1930 <- subset(ohio1930, county != 129)

#SUBSET to White males between 18,22 and higrade < 20 which would be graduate school
edu <- subset(edu, sex == "Male")
edu <- subset(edu, between(age, 18, 22))
edu <- subset(edu, higrade < 20)
edu$higrade <- edu$higrade - 3

#Remove all values less than 1
edu <- edu[which(edu$higrade > 0),]


#THIS IS FINAL 1940 EDU 
write.csv(edu, "ohio1940edu.csv")


newEdu <- edu %>%
  group_by(county) %>%
  summarise(avg = mean(higrade))

plot(ohio1930$percent_farm_tractor, newEdu$avg, xlab = "% 1930", ylab ="Avg years education", main = "Ohio")


plotSrc <- merge(newEdu, ohio1930, by = "county")

#LOESS REGRESSION on EDU vs % Tractor
ggplot(plotSrc,aes(x = percent_farm_tractor, y = avg)) + geom_point()  + geom_smooth() + xlab("1930's %") + ylab("1940's Edu") + ggtitle("Ohio")
#identify(ohio1930$percent_farm_tractor, newEdu$avg)
```

##Transform ohio edu response on Log Scale 
### Ohio Mean Model

```{r}
plotSrc$logAvg <- log(plotSrc$avg)

ohEduMod <- lm(logAvg ~ percent_farm_tractor, data = plotSrc)
plot(plotSrc$percent_farm_tractor, plotSrc$logAvg, main = "Log-Linear")
abline(ohEduMod, col = "red")
summary(ohEduMod)
```


### Ohio - Median Plots
```{r}


eduMedian <- edu %>%
  group_by(county) %>%
  summarise(med = median(higrade))

plot(jitter(ohio1930$percent_farm_tractor), jitter(eduMedian$med), xlab = "% 1930", ylab ="Median years education", main = "Ohio- Median")
eduMedian$logMed <- log(eduMedian$med)

plot(jitter(ohio1930$percent_farm_tractor), jitter(eduMedian$logMed), xlab = "% 1930", ylab ="Median years education", main = "Ohio- LogMedian")

```

### Ohio - Median Model
```{r}
plotSrc_med <- merge(eduMedian, ohio1930, by = "county")
ohEduMod_med <- lm(logMed ~ percent_farm_tractor, data = plotSrc_med)
summary(ohEduMod_med)

oh_med_df <- as.data.frame(resid(ohEduMod_med))
oh_med_df$county <- plotSrc_med$name
colnames(oh_med_df) <- c( "residual", "county")

```

### Ohio Median Model County Residuals
```{r}
oh_med_df

```

## OH EDU VS % SSE
```{r}
#Calculate SSE
pred_ohEdu <- predict(ohEduMod, plotSrc)
#Undo log transformation
pred_ohEdu <- exp(pred_ohEdu)-1
```


##Sum of Square error
```{r}
mean((plotSrc$avg - pred_ohEdu)^2)


#Residuals
plotSrc$modResid <- resid(ohEduMod)
plotSrc$fitted <- predict(ohEduMod)
ggplot(plotSrc, aes(x = fitted, y = modResid)) + geom_point() + geom_hline(yintercept = 0, colour = "red") + ggtitle(label = "OH Edu vs Tractor Residuals")
```


## VAM 1947 vs % 1940
```{r}

#SCATTER DIAGRAM AND REGRESSION FOR OHIO VAM VS EDU
#VAM dataset
vam_ohio <- subset(vam, state == "Ohio")
#write.csv(vam_ohio, "vamOhio.csv")

#VAM in 1947 vs % Tractors 1940
vam_ohio <- subset(vam_ohio, year == 1947)
ohio1940 <- subset(df40, stateAbb == "OH")
ohio1930 <- subset(df30, stateAbb == "OH")
ohio1940$county <- ohio1940$county * 10

plot(ohio1940$percent_farm_tractor, vam_ohio$VAM, xlab = "% 1940", ylab = "VAM 1947", ylim = c(0,1000), main = "Ohio VAM vs % Tractor")
abline(lm(vam_ohio$VAM ~ ohio1940$percent_farm_tractor))

```

## VAM 1958 vs % 1950
```{r}
ohio1950 <- subset(df50,stateAbb == "OH" )
vam_ohio1957 <- subset(vam, year == 1958 & state == "Ohio")
ohVam_mod <- lm(vam_ohio1957$VAM ~ ohio1950$percent_farm_tractor)
summary(ohVam_mod)

plot(ohio1950$percent_farm_tractor, vam_ohio1957$VAM, main = "Ohio - 1957 VAM vs 1950 %")
abline(ohVam_mod, col = "red")
```

## VAM 1947 vs % 1930
```{r}
vam_ohio1947 <- subset(vam, year == 1947 & state == "Ohio")
ohVam_mod1947 <- lm(vam_ohio1947$VAM ~ ohio1930$percent_farm_tractor)
summary(ohVam_mod1947)

plot(ohio1930$percent_farm_tractor, vam_ohio1947$VAM, main = "Ohio - 1947 VAM vs 1930 %")
abline(ohVam_mod1947, col = "red")
```

------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

```{r}
# NEW DATASET! -> EDU 1940 vs % Tractor 1930
oth_state <- read.dta13("cen1940A.dta")
#States:
# Michigan, Wisconsin, Indiana, Illinois, Iowa, Minnesota
```


------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

#MICHIGAN
```{r}


"use `stateicp` as locations
age between 18,22
higrade < 20"


#Michigan Edu
mich_edu <- eduDataFunc("Michigan")

#Michigan Tractor
mich_tract <- tractDataFunc("MI")

plot(mich_tract$percent_farm_tractor, mich_edu$avg, xlab = "1930 % Tractor", ylab = "1940 Avg. Edu", main = "Michigan")

michPlotSrc <- merge(mich_edu, mich_tract, by = "county")

#LOESS REGRESSION on EDU vs % Tractor
ggplot(michPlotSrc,aes(x = percent_farm_tractor, y = avg)) + geom_point()  + geom_smooth() + xlab("1930's % Tractor") + ylab("1940 Avg. Edu") + ggtitle("Michigan")
```


#Transform response on Log Scale
```{r}
michPlotSrc$logAvg <- log(michPlotSrc$avg)

plot(michPlotSrc$percent_farm_tractor, michPlotSrc$logAvg, main = "Log-Linear")
michEduMod <- lm(logAvg ~ percent_farm_tractor, data = michPlotSrc)
summary(michEduMod)

#Residuals
michPlotSrc$modResid <- resid(michEduMod)
michPlotSrc$fitted <- predict(michEduMod)
ggplot(michPlotSrc, aes(x = fitted, y = modResid)) + geom_point() + geom_hline(yintercept = 0, colour = "red") + ggtitle(label = "MI Edu vs Tractor Residuals")
```

#Michigan - Median
```{r}

michMedian <- eduMedianFunc("Michigan")

plot(jitter(mich_tract$percent_farm_tractor), jitter(michMedian$med), xlab = "% 1930", ylab ="Median years education", main = "Mich- Median")
michMedian$logMed <- log(michMedian$med)

plot(jitter(mich_tract$percent_farm_tractor), jitter(michMedian$logMed), xlab = "% 1930", ylab ="Median years education", main = "Mich- LogMedian")
```

### Michigan - Median Model
```{r}
plotSrc_med_mich <- merge(michMedian, mich_tract, by = "county")
michEduMod_med <- lm(logMed ~ percent_farm_tractor, data = plotSrc_med_mich)
summary(michEduMod_med)

mi_med_df <- as.data.frame(resid(michEduMod_med))
mi_med_df$county <- plotSrc_med_mich$name
colnames(mi_med_df) <- c( "residual", "county")

```

### Michigan Median Model County Residuals
```{r}
mi_med_df
```




#SCATTER DIAGRAM AND REGRESSION FOR MICHIGAN VAM VS EDU
```{r}
mich_vam <- subset(vam, state == "Michigan" & year == 1947)
plot(mich_tract$percent_farm_tractor, mich_vam$VAM, xlab = "% 1940", ylab = "VAM 1947", ylim = c(0,1000), main = "Michigan VAM vs % Tractor")
abline(lm(mich_vam$VAM ~ mich_tract$percent_farm_tractor))

outlierMich <- michPlotSrc[44,]
outlierMich
```


## MI EDU VS % SSE
```{r}
#Calculate SSE
pred_michEdu <- predict(michEduMod, michPlotSrc)
#Undo log transformation
pred_michEdu <- exp(pred_michEdu)-1

mean((michPlotSrc$avg - pred_michEdu)^2)
```

## VAM 1958 vs % 1950
```{r}
mich1950 <- subset(df50,stateAbb == "MI" )
mich1930 <- subset(df30,stateAbb == "MI" )
vam_mich1957 <- subset(vam, year == 1958 & state == "Michigan")
michVam_mod <- lm(vam_mich1957$VAM ~ mich1950$percent_farm_tractor)
summary(michVam_mod)

plot(mich1950$percent_farm_tractor, vam_mich1957$VAM, main = "Michigan - 1957 VAM vs 1950 %")
abline(michVam_mod, col = "red")
```

## VAM 1947 vs % 1930
```{r}
vam_mich1947 <- subset(vam, year == 1947 & state == "Michigan")
michVam_mod1947 <- lm(vam_mich1947$VAM ~ mich1930$percent_farm_tractor)
summary(michVam_mod1947)

plot(mich1930$percent_farm_tractor, vam_mich1947$VAM, main = "Michigan - 1947 VAM vs 1930 %")
abline(michVam_mod1947, col = "red")
```

-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

#WISCONSIN

```{r}

#Wisconsin Edu
wis_edu <- eduDataFunc("Wisconsin")

#Wisconsin Tractor
wis_tract <- tractDataFunc("WI")

plot(wis_tract$percent_farm_tractor, wis_edu$avg, xlab = "1930 % Tractor", ylab = "1940 Avg. Edu", main = "Wisconsin")
plot(log(wis_tract$percent_farm_tractor), wis_edu$avg)

wisPlotSrc <- merge(wis_edu, wis_tract, by = "county")

#LOESS REGRESSION on EDU vs % Tractor
ggplot(wisPlotSrc,aes(x = percent_farm_tractor, y = avg)) + geom_point()  + geom_smooth() + xlab("1930's % Tractor") + ylab("1940 Avg. Edu") + ggtitle("Wisconsin")



#Transform response on Log Scale
wisPlotSrc$logAvg <- log(wisPlotSrc$avg)
wisPlotSrc$logX <- log(wisPlotSrc$percent_farm_tractor)

plot(wisPlotSrc$percent_farm_tractor, wisPlotSrc$logAvg, main = "Log-Linear")
wisEduMod <- lm(logAvg ~ percent_farm_tractor, data = wisPlotSrc)
summary(wisEduMod)

#Residuals
wisPlotSrc$modResid <- resid(wisEduMod)
wisPlotSrc$fitted <- predict(wisEduMod)
ggplot(plotSrc, aes(x = fitted, y = modResid)) + geom_point() + geom_hline(yintercept = 0, colour = "red") + ggtitle(label = "WI Edu vs Tractor Residuals")


#SCATTER DIAGRAM AND REGRESSION FOR WISCONSIN VAM VS EDU
wis_vam <- subset(vam, state == "Wisconsin" & year == 1947)
wis_vam[is.na(wis_vam)] <- 0
plot(wis_tract$percent_farm_tractor, wis_vam$VAM, xlab = "% 1940", ylab = "VAM 1947", ylim = c(0,1000), main = "Wisconsin VAM vs % Tractor")
abline(lm(wis_vam$VAM ~ wis_tract$percent_farm_tractor))
```


## WI EDU VS % SSE
```{r}
#Calculate SSE
pred_wisEdu <- predict(wisEduMod, wisPlotSrc)
#Undo log transformation
pred_wisEdu <- exp(pred_wisEdu)-1

mean((wisPlotSrc$avg - pred_wisEdu)^2)
```


## VAM 1958 vs % 1950
```{r}
wis1950 <- subset(df50,stateAbb == "WI" )
wis1930 <- subset(df30,stateAbb == "WI" )
vam_wis1957 <- subset(vam, year == 1958 & state == "Wisconsin")
wisVam_mod <- lm(vam_wis1957$VAM ~ wis1950$percent_farm_tractor)
summary(wisVam_mod)

plot(wis1950$percent_farm_tractor, vam_wis1957$VAM, main = "Wisconsin - 1957 VAM vs 1950 %")
abline(wisVam_mod, col = "red")
```

## VAM 1947 vs % 1930
```{r}
vam_wis1947 <- subset(vam, year == 1947 & state == "Wisconsin")
wisVam_mod1947 <- lm(vam_wis1947$VAM ~ wis1930$percent_farm_tractor)
summary(wisVam_mod1947)

plot(wis1930$percent_farm_tractor, vam_wis1947$VAM, main = "Wisconsin - 1947 VAM vs 1930 %")
abline(wisVam_mod1947, col = "red")
```

-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

# INDIANA
```{r}

#Indiana Edu
in_edu <- eduDataFunc("Indiana")

#Indiana Tractor
in_tract <- tractDataFunc("IN")

plot(in_tract$percent_farm_tractor, in_edu$avg, xlab = "1930 % Tractor", ylab = "1940 Avg. Edu", main = "Indiana")

inPlotSrc <- merge(in_edu, in_tract, by = "county")

#LOESS REGRESSION on EDU vs % Tractor
ggplot(inPlotSrc,aes(x = percent_farm_tractor, y = avg)) + geom_point()  + geom_smooth() + xlab("1930's % Tractor") + ylab("1940 Avg. Edu") + ggtitle("Indiana")
```


#Transform response on Log Scale
```{r}
inPlotSrc$logAvg <- log(inPlotSrc$avg)

plot(inPlotSrc$percent_farm_tractor, inPlotSrc$logAvg, main = "Log-Linear")
inEduMod <- lm(logAvg ~ percent_farm_tractor, data = inPlotSrc)
summary(inEduMod)

#Residuals
inPlotSrc$modResid <- resid(inEduMod)
inPlotSrc$fitted <- predict(inEduMod)
ggplot(inPlotSrc, aes(x = fitted, y = modResid)) + geom_point() + geom_hline(yintercept = 0, colour = "red") + ggtitle(label = "IN Edu vs Tractor Residuals")
```

#SCATTER DIAGRAM AND REGRESSION FOR Indiana VAM VS EDU
```{r}
in_vam <- subset(vam, state == "Indiana" & year == 1947)
rownames(in_vam) <- 1:nrow(in_vam)
in_vam <- in_vam[-75,]
rownames(in_vam) <- 1:nrow(in_vam)
in_vam[is.na(in_vam)] <- 0
plot(in_tract$percent_farm_tractor, in_vam$VAM, xlab = "% 1940", ylab = "VAM 1947", ylim = c(0,1000), main = "Indiana VAM vs % Tractor")
abline(lm(in_vam$VAM ~ in_tract$percent_farm_tractor))
```

#Indiana - Median
```{r}

inMedian <- eduMedianFunc("Indiana")

plot(jitter(in_tract$percent_farm_tractor), jitter(inMedian$med), xlab = "% 1930", ylab ="Median years education", main = "Indiana- Median")
inMedian$logMed <- log(inMedian$med)

plot(jitter(in_tract$percent_farm_tractor), jitter(inMedian$logMed), xlab = "% 1930", ylab ="Median years education", main = "Indiana- LogMedian")
```



### Indiana - Median Model
```{r}
plotSrc_med_in <- merge(inMedian, in_tract, by = "county")
inEduMod_med <- lm(logMed ~ percent_farm_tractor, data = plotSrc_med_in)
summary(inEduMod_med)

in_med_df <- as.data.frame(resid(inEduMod_med))
in_med_df$county <- plotSrc_med_in$name
colnames(in_med_df) <- c( "residual", "county")

```


### Indiana Median Model County Residuals
```{r}
in_med_df
```
## IN EDU VS % SSE
```{r}
#Calculate SSE
pred_inEdu <- predict(inEduMod, inPlotSrc)
#Undo log transformation
pred_inEdu <- exp(pred_inEdu)-1

mean((inPlotSrc$avg - pred_inEdu)^2)
```

## VAM 1958 vs % 1950
```{r}
in1950 <- subset(df50,stateAbb == "IN" )
in1930 <- subset(df30,stateAbb == "IN" )
vam_in1957 <- subset(vam, year == 1958 & state == "Indiana")
inVam_mod <- lm(vam_in1957$VAM ~ in1950$percent_farm_tractor)
summary(inVam_mod)

plot(in1950$percent_farm_tractor, vam_in1957$VAM, main = "Indiana - 1957 VAM vs 1950 %")
abline(inVam_mod, col = "red")
```

## VAM 1947 vs % 1930
```{r}
vam_in1947 <- subset(vam, year == 1947 & state == "Indiana")
vam_in1947 <- vam_in1947[-75,]

inVam_mod1947 <- lm(vam_in1947$VAM ~ in1930$percent_farm_tractor)
summary(inVam_mod1947)

plot(in1930$percent_farm_tractor, vam_in1947$VAM, main = "Indiana - 1947 VAM vs 1930 %")
abline(inVam_mod1947, col = "red")
```
-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

#Illinois

```{r}
#Illinois Edu
il_edu <- eduDataFunc("Illinois")

il_edu_med <- eduMedianFunc("Illinois")

#Illinois Tractor
il_tract <- tractDataFunc("IL")

plot(il_tract$percent_farm_tractor, il_edu$avg, xlab = "1930 % Tractor", ylab = "1940 Avg. Edu", main = "Illinois")

plot(il_tract$percent_farm_tractor, il_edu_med$med, xlab = "1930 % Tractor", ylab = "1940 Med. Edu", main = "Illinois")

ilPlotSrc <- merge(il_edu, il_tract, by = "county")
```
#LOESS REGRESSION on EDU vs % Tractor
```{r}
ggplot(ilPlotSrc,aes(x = percent_farm_tractor, y = avg)) + geom_point()  + geom_smooth() + xlab("1930's % Tractor") + ylab("1940 Avg. Edu") + ggtitle("Illinois")
```

#Transform response on Log Scale
```{r}
ilPlotSrc$logAvg <- log(ilPlotSrc$avg)

plot(ilPlotSrc$percent_farm_tractor, ilPlotSrc$logAvg, main = "Log-Linear")
ilEduMod <- lm(logAvg ~ percent_farm_tractor, data = ilPlotSrc)
summary(ilEduMod)
```

#Residuals - Mean Model
```{r}
ilPlotSrc$modResid <- resid(ilEduMod)
ilPlotSrc$fitted <- predict(ilEduMod)
ggplot(ilPlotSrc, aes(x = fitted, y = modResid)) + geom_point() + geom_hline(yintercept = 0, colour = "red") + ggtitle(label = "IL Edu vs Tractor Residuals")
```


#SCATTER DIAGRAM AND REGRESSION FOR Illinois VAM VS EDU
```{r}
il_vam <- subset(vam, state == "Illinois" & year == 1947)
rownames(il_vam) <- 1:nrow(il_vam)
il_vam[is.na(il_vam)] <- 0
plot(il_tract$percent_farm_tractor, il_vam$VAM, xlab = "% 1940", ylab = "VAM 1947", ylim = c(0,1000), main = "Illinois VAM vs % Tractor")
abline(lm(il_vam$VAM ~ il_tract$percent_farm_tractor))

#Identify outliers
 #outlier <- ilPlotSrc[c(35,54),]
 #outlier
```

#Illinois - Median
```{r}

ilMedian <- eduMedianFunc("Illinois")

plot(jitter(il_tract$percent_farm_tractor), jitter(ilMedian$med), xlab = "% 1930", ylab ="Median years education", main = "Illinois- Median")
ilMedian$logMed <- log(ilMedian$med)

plot(jitter(il_tract$percent_farm_tractor), jitter(ilMedian$logMed), xlab = "% 1930", ylab ="Median years education", main = "Illinois - LogMedian")
```






### Illinois - Median Model
```{r}
plotSrc_med_il <- merge(ilMedian, il_tract, by = "county")
ilEduMod_med <- lm(logMed ~ percent_farm_tractor, data = plotSrc_med_il)
summary(ilEduMod_med)

il_med_df <- as.data.frame(resid(ilEduMod_med))
il_med_df$county <- plotSrc_med_il$name
colnames(il_med_df) <- c( "residual", "county")

```

### Illinois Median Model County Residuals
```{r}
il_med_df
```


## IL EDU VS % SSE
```{r}
#Calculate SSE
pred_ilEdu <- predict(ilEduMod, ilPlotSrc)
#Undo log transformation
pred_ilEdu <- exp(pred_ilEdu)-1

mean((ilPlotSrc$avg - pred_ilEdu)^2)
```


## VAM 1958 vs % 1950
```{r}
il1950 <- subset(df50,stateAbb == "IL" )
vam_il1957 <- subset(vam, year == 1958 & state == "Illinois")
ilVam_mod <- lm(vam_il1957$VAM ~ il1950$percent_farm_tractor)
summary(ilVam_mod)

plot(il1950$percent_farm_tractor, vam_il1957$VAM, main = "Illinois - 1957 VAM vs 1950 %")
abline(ilVam_mod, col = "red")
```

## VAM 1947 vs % 1930
```{r}
il1930 <- subset(df30, stateAbb == "IL")
ilVam_mod1947 <- lm(il_vam$VAM ~ il1930$percent_farm_tractor)
summary(ilVam_mod1947)

plot(il1930$percent_farm_tractor, il_vam$VAM, main = "Illinois - 1947 VAM vs 1930 %")
abline(ilVam_mod1947, col = "red")
```

-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

#IOWA

```{r}
#Iowa Edu
ia_edu <- eduDataFunc("Iowa")

#Iowa Tractor
ia_tract <- tractDataFunc("IA")

plot(ia_tract$percent_farm_tractor, ia_edu$avg, xlab = "1930 % Tractor", ylab = "1940 Avg. Edu", main = "Iowa")

iaPlotSrc <- merge(ia_edu, ia_tract, by = "county")
```

#LOESS REGRESSION on EDU vs % Tractor
```{r}
ggplot(iaPlotSrc,aes(x = percent_farm_tractor, y = avg)) + geom_point()  + geom_smooth() + xlab("1930's % Tractor") + ylab("1940 Avg. Edu") + ggtitle("Iowa")

#Transform response on Log Scale
iaPlotSrc$logAvg <- log(iaPlotSrc$avg)

plot(iaPlotSrc$percent_farm_tractor, iaPlotSrc$logAvg, main = "Log-Linear")
iaEduMod <- lm(logAvg ~ percent_farm_tractor, data = iaPlotSrc)
summary(iaEduMod)
```

#Residuals
```{r}
iaPlotSrc$modResid <- resid(iaEduMod)
iaPlotSrc$fitted <- predict(iaEduMod)
ggplot(iaPlotSrc, aes(x = fitted, y = modResid)) + geom_point() + geom_hline(yintercept = 0, colour = "red") + ggtitle(label = "IA Edu vs Tractor Residuals")
```

#SCATTER DIAGRAM AND REGRESSION FOR Indiana VAM VS EDU
```{r}
ia_vam <- subset(vam, state == "Iowa" & year == 1947)
rownames(ia_vam) <- 1:nrow(ia_vam)
ia_vam[is.na(ia_vam)] <- 0
plot(ia_tract$percent_farm_tractor, ia_vam$VAM, xlab = "% 1940", ylab = "VAM 1947", ylim = c(0,1000), main = "Iowa VAM vs % Tractor")
abline(lm(ia_vam$VAM ~ ia_tract$percent_farm_tractor))
```

## IA EDU VS % Sum of Squares Error
```{r}
#Calculate SSE
pred_iaEdu <- predict(iaEduMod, iaPlotSrc)
#Undo log transformation
pred_iaEdu <- exp(pred_iaEdu)-1

mean((iaPlotSrc$avg - pred_iaEdu)^2)
```

## VAM 1958 vs % 1950
```{r}
ia1950 <- subset(df50,stateAbb == "IA" )
vam_ia1957 <- subset(vam, year == 1958 & state == "Iowa")
iaVam_mod <- lm(vam_ia1957$VAM ~ ia1950$percent_farm_tractor)
summary(iaVam_mod)
plot(ia1950$percent_farm_tractor, vam_ia1957$VAM, main = "Iowa - 1957 VAM vs 1950 %")
abline(iaVam_mod, col = "red")

```

## VAM 1947 vs % 1930
```{r}
ia1930 <- subset(df30, stateAbb == "IA")
iaVam_mod1947 <- lm(ia_vam$VAM ~ ia1930$percent_farm_tractor)
summary(iaVam_mod1947)

plot(ia1930$percent_farm_tractor, ia_vam$VAM, main = "Iowa - 1947 VAM vs 1930 %")
abline(iaVam_mod1947, col = "red")
```

#Iowa - Median
```{r}
iaMedian <- eduMedianFunc("Iowa")

plot(jitter(ia_tract$percent_farm_tractor), jitter(iaMedian$med), xlab = "% 1930", ylab ="Median years education", main = "Iowa - Median")
iaMedian$logMed <- log(iaMedian$med)

plot(jitter(ia_tract$percent_farm_tractor), jitter(iaMedian$logMed), xlab = "% 1930", ylab ="Median years education", main = "Iowa - LogMedian")
```

### Iowa - Median Model
```{r}
plotSrc_med_ia <- merge(iaMedian, ia_tract, by = "county")
iaEduMod_med <- lm(logMed ~ percent_farm_tractor, data = plotSrc_med_ia)
summary(iaEduMod_med)

ia_med_df <- as.data.frame(resid(iaEduMod_med))
ia_med_df$county <- plotSrc_med_ia$name
colnames(ia_med_df) <- c( "residual", "county")

```

### Iowa Median Model County Residuals
```{r}
ia_med_df
```

-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

